The very dependent recursive structure of iterated parametricity in indexed form

### 2025-12-29 paper/paper.tex
paper: s/tabular/tabularx/
paper: tweak experimental table

The internalisation of parametricity in unary indexed form was studied by \cite{Moulin,Bernardy} directly leading to a type theory of iterated parametricity. Such type theory could be interpreted in a ``unary'' form of cubical set~\cite{bernardy15}, thus implicitly in fibered form since, in a binary cubical set (resp. in a unary cubical set and simplicial set as well), a set of segments over two points, that is a relevant relation, is given by first and second projection rather than as a function to $\Set$ (resp. a family is given as a \emph{total space} projected over a set of points $A:\Set$ rather than a family of type $A \rightarrow \Set$. This connection between iterated binary parametricity in fibered form and cubical sets includes an identification of degeneracies and reflexivity proofs, which can also be found in \cite{johann17}. In particular, it became clear at this point that iterated binary parametricity (without reflexivity) produces semi-cubical sets while it was also observed\footnote{The authors learned it from Hugo Moeneclaye and Thorsten Altenkirch.} that iterated unary parametricity produces augmented semi-simplicial sets.

\begin{tabular}{lllllll}
                            &                              &                                 &                               & structure                          & kind of semantics   \\
  \hline
  \multirow{6}{1cm}{unary}  & \multirow{3}{1.5cm}{indexed} & \multirow{2}{2cm}{non-iterated} & proof-irrelevant              & predicate                          & reducibility        \\
                            &                              &                                 & relevant                      & family                             &                     \\
                            &                              & \multicolumn{2}{l}{iterated}    & \emph{this work}              &                                                          \\
                            & \multirow{3}{1.5cm}{fibered} & \multirow{2}{2cm}{non-iterated} & proof-irrelevant              & subobject                          &                     \\
                            &                              &                                 & relevant                      & fibration                          &                     \\
                            &                              & \multicolumn{2}{l}{iterated}    & augmented semi-simplicial set &                                                          \\
  \hline
  \multirow{6}{1cm}{binary} & \multirow{3}{1.5cm}{indexed} & \multirow{2}{2cm}{non-iterated} & proof-irrelevant              & relation                           & logical relation    \\
                            &                              &                                 & relevant                      & correspondence (relevant relation) &                     \\
                            &                              & \multicolumn{2}{l}{iterated}    & \emph{this work}              &                                                          \\
                            & \multirow{3}{1.5cm}{fibered} & \multirow{2}{2cm}{non-iterated} & proof-irrelevant              & fibered relation ?                 &                     \\
                            &                              &                                 & relevant                      & span                               &                   & \\
                            &                              & \multicolumn{2}{l}{iterated}    & semi-cubical set              &                                                          \\
  \hline
\end{tabular}

>>>>

The internalisation of parametricity in unary indexed form was studied by \cite{Moulin,Bernardy} directly leading to a type theory of iterated parametricity. Such type theory could be interpreted in a ``unary'' form of cubical set~\cite{bernardy15}, thus implicitly in fibered form since, in a binary cubical set (resp. in a unary cubical set and simplicial set as well), a set of segments over two points, that is a relevant relation, is given by first and second projection rather than as a function to $\Set$ (resp. a family is given as a \emph{total space} projected over a set of points $A:\Set$ rather than a family of type $A \rightarrow \Set$. This connection between iterated binary parametricity in fibered form and cubical sets includes an identification of degeneracies and reflexivity proofs, which can also be found in \cite{johann17}. In particular, it became clear at this point that iterated binary parametricity (without reflexivity) produces semi-cubical sets while it was also observed\footnote{The authors learned it from Hugo Moeneclaey and Thorsten Altenkirch.} that iterated unary parametricity produces augmented semi-simplicial sets.

\begin{tabularx}{\linewidth}{lYcYYYl}
  \toprule
  arity                     & formalisation style          & iteration                       & relevance                     & structure          & name of semantics   \\
  \midrule
  \multirow{6}{1cm}{unary}  & \multirow{3}{1.5cm}{indexed} & \multirow{2}{2cm}{non-iterated} & proof-irrelevant              & predicate          & reducibility        \\
                            &                              &                                 & relevant                      & family             &                     \\
                            &                              & \multicolumn{2}{l}{iterated}    & \emph{this work}              &                                          \\
                            & \multirow{3}{1.5cm}{fibered} & \multirow{2}{2cm}{non-iterated} & proof-irrelevant              & subobject          &                     \\
                            &                              &                                 & relevant                      & fibration          &                     \\
                            &                              & \multicolumn{2}{l}{iterated}    & augmented semi-simplicial set &                                          \\
  \midrule
  \multirow{6}{1cm}{binary} & \multirow{3}{1.5cm}{indexed} & \multirow{2}{2cm}{non-iterated} & proof-irrelevant              & relation           & logical relation    \\
                            &                              &                                 & relevant                      & correspondence     &                     \\
                            &                              & \multicolumn{2}{l}{iterated}    & \emph{this work}              &                                          \\
                            & \multirow{3}{1.5cm}{fibered} & \multirow{2}{2cm}{non-iterated} & proof-irrelevant              & fibered relation ? &                     \\
                            &                              &                                 & relevant                      & span               &                   & \\
                            &                              & \multicolumn{2}{l}{iterated}    & semi-cubical set              &                                          \\
  \bottomrule
\end{tabularx}
